### 2016-12-11 doc/arxivize.tex
added doc

\pdfoutput=1

\documentclass[letterpaper]{article}
\usepackage{hyperref}
\hypersetup{
  pdfinfo={
    Title={Entropy-SGD: Biasing Gradient Descent Into Wide Valleysre},
    Author={Pratik Chaudhari, Anna Choromanska, Stefano Soatto, Yann LeCun},
    Keywords={deep learning, gradient descent, optimization, machine learning, neural networks}
  }
}
\usepackage{pdfpages}
\usepackage{grffile}

\begin{document}
\includepdf[pages=1-last]{chaudhari.choromanska.ea.iclr17.pdf}
\end{document}