### 2017-01-02 doc/rebuttal.txt
rebuttal v1.0

AC rebuttal
==========

We are happy to provide clarifications to the questions posed. We enclose them below.

"mostly reinventing SVI"
SVI computes the posterior distribution of hidden variables given data by maximizing the ELBO. Please note that we do not have hidden variables in our formulation; "x" are the variables of our objective f(x), not data. We are interested in maximizing the free energy F(x,\gamma) as defined in Eqn. 4. This is the *free energy of the energy landscape* (defined via a Gibbs distribution). This is thus unrelated to marginal likelihood or variational ELBO where one can impose entropic constraints on the posterior distribution of the *hidden variables* via a prior.

"I suspect it'd become clear that you are maximizing the marginal likelihood"
Please note that we are not maximizing the marginal likelihood, as discussed above.

"how is this better than SGLD", "bizarre that you compare against Adam instead of SGLD"
SGLD is an MCMC algorithm that draws samples from a given distribution. If the step-size goes to zero slowly enough, akin to all MCMC algorithms, it converges to the maximum of the likelihood in exponentially long time-scales (see algorithms like SGHMC [CFG14], Santa [CCG15] etc. that train Bayesian neural networks using MCMC algorithms). Note that SGLD does not optimize the marginal likelihood because there is no notion of hidden variables in vanilla SGLD. We provide a brief review of SGLD in Appendix A of our paper.

We do not know of any results in the literature that train large deep networks such as the one used for CIFAR to get competitive error rates using SGLD. We would like to emphasize that Entropy-SGD simply uses MCMC sampling to estimate Eqn. 7 but it is unrelated to SGLD otherwise.

On the other hand, Adam is an algorithm for computing the maximum of a the likelihood of data given parameters or (equivalently, minimize the loss function). Entropy-SGD is an algorithm designed for minimizing the loss function f(x), in particular, it is not an MCMC algorithm that draws samples from the likelihood. It however does not explicitly do so and instead maximizes the local entropy. We therefore compare Entropy-SGD with state-of-the-art algorithms for training deep networks like Adam and SGD.

"frustrating that you discuss free energy and entropy without precise definitions"
Local entropy (local free energy) is formally defined in Def. 1 (Eqn. 4) but it is already introduced on pg. 2 in the Introduction. The discussion towards the end of Sec. 3 (pg. 5, first line) defines the classical entropy of a Gibbs distribution. The beginning of Sec. 3 based on Fig. 2 is intended to explain things to the reader at an intuitive level before proceeding to the formal definitions which are nevertheless present.

Free energy and classical entropy are related by the informal description "free energy = internal energy - temperature x entropy". Formally, the relation is
    F(\beta) = U(\beta) - \frac{1}{\beta} S(\beta)
where the log-partition function (free energy) is defined as F(\beta) = -\beta^{-1} \log Z(\beta), the internal energy is U(\beta) = \partial_\beta (\beta F(\beta)) and S(\beta) = \beta^2 \partial_\beta (F(\beta)) is the classical entropy.

Note that the above equation is about the entire Gibbs distribution, in our work we define a "local free energy" F(x, \beta) via the modified Gibbs distribution in Eqn. 3. We can add this discussion to the paper.

[CFG14] Chen, Tianqi, Emily B. Fox, and Carlos Guestrin. "Stochastic Gradient Hamiltonian Monte Carlo." ICML. 2014.
[CCF15] Chen, Changyou, et al. "Bridging the gap between stochastic gradient mcmc and stochastic optimization." arXiv:1512.07962 (2015).

----------------------------------------------------------------------

Title: You cannot do SVI on the network weights unless you choose a “moving prior”, a feature of our scheme.

To address the objection 1) let \Xi denote the dataset, z  the weights and x the parameters of a variational distribution q_x(z). The ELBO can then be written as
(i) \log p(\Xi)
        \geq E_{z \sim q_x(z)} [\log p(\Xi | z)] - KL(q_x(z) || p(z))
        = E_{z \sim q_x(z)} [-f(z)] - KL(q_x(z) || p(z));
and maximized wrt x. The distribution p(z) is a fixed (parameter-free) prior, which one has to postulate.

On the other hand, Eqn. 4 in the paper can be used to write the log of local entropy, where f = -\logp(\Xi | z):
(ii) \log F(x,\gamma)
        = \log \int_{z \in Z} e^{-f(z; \Xi) - \gamma/2 |x-z|^2} dz.
We are unaware of any way to choose a prior p(z) and a variational family q_x(z) that that makes (i) resemble (ii), and therefore interpret our method as “integrating out […] the posterior over neural network parameters.” The only way we can do so is to pick a specific “prior" that depends on the parameter x (hence, not really a prior). For instance, one could choose a uniform variational family and a Gaussian “prior" with mean x to make (ii) resemble ELBO. In this case p(z) would not be fixed, but it would “move” along with the current iterate x. This "moving prior" is a crucial feature of our proposed algorithm.

Concerning the relation to SGLD, consider Belief Propagation (BP). Our proposed algorithm relates to the "focusing-Belief Propagation" variant (fBP), rather than the standard one [BBC16]. The difference between BP and fBP is analogous to that between SLGD and entropy-SGD: the latter operates on a transformation of the landscape of the former, exploiting local entropic effects. This difference is crucial and indeed related to the "moving prior" of the previous discussion; plain SGLD (or BP) can only trade accuracy for entropy via the temperature parameter, which does not allow to directly use any geometric information on the landscape and does not help with narrow minima.

In view of your comment 2), we also implemented SGLD for LeNet on MNIST and All-CNN-BN on CIFAR and will add the following results to our paper: After a hyper-parameter search, the best rwe obtained were a test error of LeNet on 0.63 \pm 0.1% on MNIST after 300 epochs and 9.89 \pm 0.11% on All-CNN-BN after 500 epochs. Disregarding the slow convergence of SGLD, its generalization error is slightly worse than the results in our paper, viz. 0.48% with Entropy-SGD on LeNet (0.51% with SGD) and 8.65% with Entropy-SGD on All-CNN-BN (8.3% with SGD). For comparison, the authors in [CCF15] report an error of 0.71% with SGLD on MNIST with a slightly larger network (0.47% with Santa), there are no results in the literature where MCMC methods perform comparably to SGD on larger networks.

[BBC16] Baldassi, Carlo, et al. "Unreasonable effectiveness of learning neural networks: From accessible states and robust ensembles to basic algorithmic schemes." PNAS (2016).
[CCF15] Chen, Changyou, et al. "Bridging the gap between stochastic gradient MCMC and stochastic optimization." arXiv:1512.07962 (2015).

-------------------------------------------------------------------------

Title: Re: Is the Hessian-vector product actually expensive?

Comment 3) above is for HVP of the loss function of a general deep network, not local entropy. Indeed, HVPs can be computed using forward differencing with two back-props and Hessian-free methods do exploit this, although to a limited degree of success [M10]. Such an approximation is susceptible to numerical errors --- especially in high dimensions [P94, M10]. One typically averages the HVP over many samples in the dataset which is expensive (for instance, the authors in [LSP93] average over a few hundred samples, which roughly translates to 5-10x one iteration of vanilla SGD). More accurate algorithms like that of [P94] also require this averaging over samples.

Practically speaking, one could argue that accuracy in the approximation of the HVP does not matter in SGD; what matters is only the local curvature at the scale of the perturbation, which needs to be chosen carefully.  In this case, indeed, approximate computation of HVP can be considered cheap.

[LSP93] LeCun, Yann, Patrice Y. Simard, and Barak Pearlmutter. "Automatic learning rate maximization by on-line estimation of the Hessian’s eigenvectors." NIPS (1993).
[P94] Pearlmutter, Barak A. "Fast exact multiplication by the Hessian." Neural computation 6.1 (1994): 147-160.
[M10] Martens, James. "Deep learning via Hessian-free optimization." ICML (2010).