### 2017-01-25 doc/arxivize.tex
uploaded to arxiv

\pdfoutput=1

\documentclass[letterpaper]{article}
\usepackage{hyperref}
\hypersetup{
  pdfinfo={
    Title={Entropy-SGD: Biasing Gradient Descent Into Wide Valleysre},
    Author={Pratik Chaudhari, Anna Choromanska, Stefano Soatto, Yann LeCun,Carlo Baldassi,Christian Borgs,Jennifer Chayes,Levent Sagun,Riccardo Zecchina},
    Keywords={deep learning, gradient descent, optimization, machine learning, neural networks,statistical physics, local entropy}
  }
}
\usepackage{pdfpages}
\usepackage{grffile}

\begin{document}
\includepdf[pages=1-last]{chaudhari.choromanska.ea.iclr17.pdf}
\end{document}